---
title: "Transcription Factor co-expression"
author: "Davide Risso, Diya Das"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    fig_width: 10
    fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=FALSE, results="markup", autodep = TRUE)

options(getClass.msg=FALSE) #get rid of annoying messages about cache until fixed internally in R
NMF::nmf.options(grid.patch=TRUE) #get rid of blank page in heatmap pdf

library(clusterExperiment)
library(NMF)
library(matrixStats)
library(igraph)

legend.col <- function(col, lev){
  opar <- par
  n <- length(col)
  bx <- par("usr")
  box.cx <- c(bx[2] + (bx[2] - bx[1]) / 1000,
              bx[2] + (bx[2] - bx[1]) / 1000 + (bx[2] - bx[1]) / 50)
  box.cy <- c(bx[3], bx[3])
  box.sy <- (bx[4] - bx[3]) / n
  xx <- rep(box.cx, each = 2)
  
  par(xpd = TRUE)
  for(i in 1:n){
    yy <- c(box.cy[1] + (box.sy * (i - 1)),
            box.cy[1] + (box.sy * (i)),
            box.cy[1] + (box.sy * (i)),
            box.cy[1] + (box.sy * (i - 1)))
    polygon(xx, yy, col = col[i], border = col[i])
  }
  par(new = TRUE)
  plot(0, 0, type = "n",
       ylim = c(min(lev), max(lev)),
       yaxt = "n", ylab = "",
       xaxt = "n", xlab = "",
       frame.plot = FALSE)
  axis(side = 4, las = 2, tick = FALSE, line = .25)
  par <- opar
}
## source: https://aurelienmadouasse.wordpress.com/2012/01/13/legend-for-a-continuous-color-scale-in-r/
```

We start from the normalized expression of  transcription factors (TFs) in the OE data.
Input include the inferred pseudo-time by Slingshot and the cluster labels by clusterExperiment.

We want to visualize the correlation of the TFs and find "modules" of co-regulated TFs. This should be visualized by a co-regulation network.

# All TFs

```{r}
expt_str <- "oeHBCregenWT"
out_dir <- file.path("../output/clust", expt_str)
viz_dir <- file.path("../output/viz", expt_str)

load(file.path(out_dir, paste0(expt_str, "_lineageData.Rda")))
load(file.path(out_dir, paste0(expt_str,"_PCA.Rda")))

nlm_tf <- nlm[intersect(unlist(read.table("../ref/ATFDB_mm_TF.txt")),rownames(nlm)),]
slm_tf <- slm[intersect(unlist(read.table("../ref/ATFDB_mm_TF.txt")), rownames(slm)),]
```

To select a reasonable threshold, we scramble the entries of the matrix and we compute the maximum correlation.

```{r}
set.seed(21421)
mm <- 0
mat <- nlm_tf;
mat <- slm_tf
for(i in seq_len(10)){
idx <- replicate(NCOL(mat), sample(seq_len(NROW(mat))))
random_mat <- matrix(mat[idx], ncol=NCOL(mat), nrow=NROW(mat))
random_cor <- cor(t(random_mat))
diag(random_cor) <- 0
mm <- max(mm, max(random_cor))
} 
print(mm)
```

# DE TFs - OneVAll

```{r}
de_NL <- read.table(paste0("../output/DE/",expt_str,"/oeHBCregenWT_NL_DE_oneVall_500.txt"), stringsAsFactors = FALSE)
de_SL <- read.table(paste0("../output/DE/",expt_str,"/oeHBCregenWT_SL_DE_oneVall_500.txt"), stringsAsFactors = FALSE)

tf_heatmap <- function(mat_tf, de, col.pal, name){
var_tf <- mat_tf[intersect(de$Feature, rownames(mat_tf)),]
cors <- cor(t(var_tf))
breakv <- c(min(var_tf), seq(0, quantile(var_tf[var_tf > 0], .98, na.rm = TRUE), length = 50), max(var_tf))

pdf(file=file.path(viz_dir, paste0(expt_str, "_DE_OneVall_HM_",name,Sys.Date(),".pdf")), height = 2,width=4)
NMF::aheatmap(var_tf, color=seqPal5, Colv=NA, Rowv=NA, annCol = data.frame(Clusters=clus.labels[colnames(var_tf)]), annColors = list(Clusters=col.pal), breaks = breakv)
dev.off()
}

tf_heatmap(nlm_tf, de=de_NL, col.pal, name="Neur")
tf_heatmap(slm_tf, de=de_SL, col.pal, name="Sus")
```

```{r}
tf_network <- function(mat_tf, de, col.pal, clus.labels, name){
  var_tf <- mat_tf[intersect(de$Feature, rownames(mat_tf)),]
  cors <- cor(t(var_tf))
  tfs <- var_tf[which(rowSums(cors>.3)>=5),]
  cor_mat <- cor(t(tfs))
  cor_mat[lower.tri(cor_mat, diag=TRUE)] <- 0
  cor_mat[cor_mat<0.3] <- 0
  
  graph <- graph.adjacency(cor_mat > 0.3, weighted=NULL, mode="upper")
  idx <- names(which(degree(graph)>=5))
  tfs <- mat_tf[idx,]
  
  cor_hm <- cor_mat <- cor(t(tfs))
  cor_mat[lower.tri(cor_mat, diag=TRUE)] <- 0
  cor_mat[cor_mat<0.3] <- 0
  
  graph <- graph.adjacency(cor_mat > 0.3, weighted=NULL, mode="upper")
  coords <- layout_(graph, with_fr())
  
  size <- apply(tfs, 1, function(x) max(tapply(x, droplevels(clus.labels), mean)))
  
  E(graph)$weight<-t(cor_mat)[t(cor_mat)>0.3]
  E(graph)$width <- E(graph)$weight*2
  V(graph)$size <- size
  names <- rownames(tfs)
  names[size < 3] <- NA
  V(graph)$label <- names

  cols_all <-apply(tfs,1, function(x) tapply(x, droplevels(clus.labels), mean))
  cols_all <- t(cols_all)
  
  par(mar=c(5.1 ,4.1 ,4.1, 2.1))
  col_ranks <- t(apply(cols_all, 1, rank))
  V(graph)$color <- "white"
  png(filename=file.path(viz_dir, paste0(expt_str,"_DE_",name,"_rankbyclus0",i,"_white_",Sys.Date(),".png")), width=8, 
      height=8, units="in", res=600, 
      pointsize=12)
  par(mar=c(5.1, 4.1, 4.1, 3))
  plot(graph, layout=coords, vertex.label.cex=.5, vertex.label.color="black")
  dev.off()
  for(i in seq_along(levels(droplevels(clus.labels)))) {
    cols = col_ranks[,i]
    clID = levels(droplevels(clus.labels))[i]
    col_ramp2 <- colorRampPalette(c("navyblue","mediumblue","dodgerblue3","aquamarine4","green4","yellowgreen","yellow"))(ncol(cols_all))
    pt_cols <- col_ramp2[cols]
    V(graph)$color <- pt_cols
    png(filename=file.path(viz_dir, paste0(expt_str,"_DE_",name,"_rankbyclus0",i,"_cl",clID, Sys.Date(), ".png")), 
        width=8, 
        height=8, units="in", res=600, 
        pointsize=12)
    par(mar=c(5.1, 4.1, 4.1, 3))
    plot(graph, layout=coords, vertex.label.cex=.5, vertex.label.color="black", main=paste("Cluster",clID, "color-coded by expression"))
    legend.col(col=col_ramp2, lev=c(1,ncol(cols_all)))
    dev.off()
  }
  
  png(file=file.path(viz_dir, paste0(expt_str,"_DE_OneVall_",name, Sys.Date(),".png")),
      width=8, 
      height=8, units="in", res=600, 
      pointsize=12)
  cols <- apply(tfs, 1, function(x) which.max(tapply(x, clus.labels, mean)))
  V(graph)$color <- col.pal[cols]
  plot(graph, layout=coords, vertex.label.cex=.5, vertex.label.color="black", main="Color-coded by cluster label")
  dev.off()
  
  write.table(data.frame(TF= rownames(tfs), Cluster_max = levels(clus.labels)[cols]), file=file.path(viz_dir, paste0(expt_str,"_DE_OneVall_TFlabels",name, Sys.Date(),".txt")), quote=F, col.names=F, row.names=F) 
}

tf_network(nlm_tf, de_NL, col.pal, nlclusters, "Neur")
tf_network(slm_tf, de_SL, col.pal, slclusters, "Sus")
```